---
title: "HITECH_Act_Case"
author: "Karl Grindal"
date: "6/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("dplyr")
library("here")
# Case 1: HITECH Security Law

# Enacted: February 17, 2009
# Enforcement: May 27, 2009

AllStateClean <- read.table(here::here("Data","Other_data","AllStateClean.txt"), sep=";")

AllStateClean$firstdate <- substr(AllStateClean$reported_date, start=1, stop=10) # Use this string for Massachusetts v New Hampshire

HITECH1 <- subset(AllStateClean, AllStateClean$year == 2005)
HITECH2 <- subset(AllStateClean, AllStateClean$year == 2006)
HITECH3 <- subset(AllStateClean, AllStateClean$year == 2007)
HITECH4 <- subset(AllStateClean, AllStateClean$year == 2008)
HITECH5 <- subset(AllStateClean, AllStateClean$year == 2009)
HITECH6 <- subset(AllStateClean, AllStateClean$year == 2010)

HITECH <- rbind(HITECH1, HITECH2, HITECH3, HITECH4, HITECH5, HITECH6)

write.csv(HITECH,here::here("Data","Other_data","HITECH.csv"))

```




```{r}

HITECH <- read.csv(here::here("Data","Other_data","HITECH.csv"))
DUNS <- read.csv(here::here("Data","Other_data","DUNS_HITECH.csv"))

NAICS <- merge(HITECH, DUNS, by="clean_name", all=TRUE)

NAICS$DUNS_2 <- substr(NAICS$Primary.NAICS.Code, start=1, stop=2)
# colSums(!is.na(NAICS)) # 1740 resolved for Duns_2 
# colSums(is.na(NAICS)) # 248 not resolved for Duns_2

248/1740 # = 14.2% are not included
1740/(1740+248) # = 87.5% of incidents are resolved

write.csv(NAICS, here::here("Data","Other_data","NAICS_Clean.csv"))

```


# Create Population Time Series for Matching with Incident Frequncy
```{r}
# Population
pop <- read.csv(here::here("Data","Other_data","populations.csv"))
pop <- pop[c(12,20,22,30,42),] # Hawaii row 12, Maine row 20, Massachusetts row 22, New Hampshire row 30, North Carolina row 34
pop

datforpop <- data.frame(seq(as.Date("2000-04-01"), by="1 month", length.out=(length(pop)-1)))
names(datforpop) <- "yearmonth"
datforpop <- format(datforpop,  "%Y/%m")
datforpop <- rbind("yearmonth",datforpop)
row.names(datforpop) <- 1:nrow(datforpop)

pop <- cbind(datforpop,t(pop))
colnames(pop) <- pop[1,]
pop <- pop[-1,]
rownames(pop) <- seq(1:nrow(pop))
pop <- as.data.frame(pop)

# Create a new column for the total population accross collecting states
pop[,2:6] <- sapply(pop[,2:6],as.numeric)
pop$totpop <- rowSums(pop[ ,c(2:6)], na.rm=TRUE)
pop

```


# Create Treatment and Control Groups
```{r}
# HIGH TECH Act regulations become effective on February 17, 2009
# Enforcement of HIGH TECH Act implement on May 27, 2009

experiment_start <- as.Date("02/17/2009", "%m/%d/%Y")-180 # Legislation H.B. 4144 becomes effective
experiment_end <- as.Date("05/27/2009", "%m/%d/%Y")+180 # moving avg uses 30 days post treatment

datfull <- data.frame(seq(as.Date(experiment_start), by="1 month", length.out=16))
names(datfull) <- "yearmonth"
datfull <- format(datfull,  "%Y/%m")
datfull

NAICS$date_formatted <- format(as.Date(NAICS$firstdate, "%Y-%m-%d"), "%Y/%m") # Alternative is "%m/%d/%Y"
NAICS_ts <- NAICS %>% 
  dplyr::group_by(NAICS$date_formatted) %>% 
  dplyr::summarise(frequency = n())

names(NAICS_ts)[1] <- "yearmonth"
NAICS_ts <- merge(NAICS_ts, pop, by="yearmonth", all.x=TRUE)
NAICS_ts$frequency[is.na(NAICS_ts$frequency)]<-0
NAICS_ts

# Look at Raw Frequency Counts

NAICS_ts <- ts(NAICS_ts$frequency, frequency = 12, start = c(2005,1))

ts.plot(NAICS_ts, main = "Breaches over time", xlim=c(2005,2011), gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Count per month")

```


# Create potential treatment and population groups
```{r}

# Format control dates into months
NAICS$date_formatted <- format(as.Date(NAICS$firstdate, "%Y-%m-%d"), "%Y/%m") # Alternative is "%m/%d/%Y"
NAICS <- NAICS[!is.na(NAICS$date_formatted),]

Health <- NAICS[grepl( "62", NAICS$DUNS_2), ]
Finance <- NAICS[grepl( "52", NAICS$DUNS_2), ]
Education <- NAICS[grepl( "61", NAICS$DUNS_2), ]
Information <- NAICS[grepl( "51", NAICS$DUNS_2), ]

Non_Health <- NAICS %>%
                   filter(DUNS_2 != 62)   

```


# Experiment 1: Create control and treatment populations (Massachusetts v New Hampshire) with Total Incidents
```{r}
treatment <-  Health # This Must Be Filled in to Work Properly!
control <-  Non_Health  # This Must Be Filled in to Work Properly!

# Format treatment dates into months
treatment$date_formatted <- format(as.Date(treatment$reported_date, "%Y-%m-%d"), "%Y/%m") # Alternative is "%m/%d/%Y"
treatment_freq <- treatment %>% 
  dplyr::group_by(treatment$date_formatted) %>% 
  dplyr::summarise(frequency = n(),)

names(treatment_freq)[1] <- "yearmonth"
treatment_freq$frequency[is.na(treatment_freq$frequency)]<-0
treatment_freq<- merge(datfull,treatment_freq, by="yearmonth", all=TRUE)

# Format control dates into months
control$date_formatted <- format(as.Date(control$reported_date, "%Y-%m-%d"), "%Y/%m") # Alternative is "%m/%d/%Y"
control_freq <- control %>% 
  dplyr::group_by(control$date_formatted) %>% 
  dplyr::summarise(frequency = n())

names(control_freq)[1] <- "yearmonth"
control_freq$frequency[is.na(control_freq$frequency)]<-0
control_freq<- merge(datfull,control_freq, by="yearmonth", all=TRUE)

treatment_ts <- ts(treatment_freq$frequency, frequency = 12, start = c(2006,8))
control_ts <- ts(control_freq$frequency, frequency = 12, start = c(2005,12))

plot.ts(treatment_ts, main = "Breaches over time", xlim=c(2006,2011), xlab = "Years", ylab = "Count per month")
plot.ts(control_ts, main = "Breaches over time", xlim=c(2006,2011), xlab = "Years", ylab = "Count per month")

ts.plot(control_ts, treatment_ts, main = "Breaches over time", xlim=c(2006,2011), gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Count per month")

```


## Create charts with breaches per million residents
```{r}
# Merge Treatment and Control Together
comb_ts <- merge(treatment_freq, control_freq, by="yearmonth", all=TRUE)

# Merge Combined Treatment and Control Together with Population Statistics
comb_ts <- merge(comb_ts, pop, by='yearmonth', all.x = TRUE)
comb_ts$frequency.x[is.na(comb_ts$frequency.x)]<-0
comb_ts$frequency.y[is.na(comb_ts$frequency.y)]<-0

comb_ts$totpop <- as.numeric(as.character(comb_ts$totpop))
comb_ts$treatpermil <- comb_ts$frequency.x/(comb_ts$totpop/1000000)
class(comb_ts$frequency.y)<-"numeric"
comb_ts$controlpermil <- comb_ts$frequency.y/(comb_ts$totpop/1000000)

treatment_tsM <- ts(comb_ts$treatpermil, frequency = 12, start = c(2006,6))
control_tsM <- ts(comb_ts$controlpermil, frequency = 12, start = c(2006,6))

plot.ts(treatment_tsM, main = "Breaches per Million over time", xlim=c(2006,2012), xlab = "Years", ylab = "Breaches per Million Resident During Month")
plot.ts(control_tsM, main = "Breaches per Million over time", xlim=c(2006,2012), xlab = "Years", ylab = "Breaches per Million Resident During Month")

ts.plot(control_tsM, treatment_tsM, main = "Breaches over time", xlim=c(2006,2012), gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Breaches per Million Resident During Month")
```

## Decompose the Control to Find Seasonal Patterns
```{r}
control_ts
controltimeseriescomponents <- decompose(control_ts)
plot(controltimeseriescomponents)
controltimeseriesseasonallyadjusted <- control_ts - controltimeseriescomponents$seasonal
plot(controltimeseriesseasonallyadjusted)

```

## Identifying and subsetting relevant dates
```{r}
# Identifying relevant dates

# HIGH TECH Act regulations become effective on February 17, 2009
# Enforcement of HIGH TECH Act implement on May 27, 2009

treatment_start <- as.Date("02/17/2009", "%m/%d/%Y") # Legislation H.B. 4144 becomes effective
treatment_start<- format(as.Date(as.character(treatment_start), origin = "1970-01-01"), "%Y/%m")

treatment_end <- as.Date("05/27/2009", "%m/%d/%Y") # post 6 months after enforcement
treatment_end<- format(as.Date(as.character(treatment_end), origin = "1970-01-01"), "%Y/%m")

pretreat <- comb_ts[(which(comb_ts$yearmonth==treatment_start)-6):which(comb_ts$yearmonth==treatment_start),]
pretreat$type <- "pretest"

posttreat <- comb_ts[which(comb_ts$yearmonth==treatment_end):(which(comb_ts$yearmonth==treatment_end)+6),]
posttreat$type <- "posttest"

mean(posttreat$treatpermil) - mean(pretreat$treatpermil)
mean(posttreat$controlpermil) - mean(pretreat$controlpermil)

treatment_range <- comb_ts[(which(comb_ts$yearmonth==treatment_start)+1):(which(comb_ts$yearmonth==treatment_end)-1),]
treatment_range$type <- "test"

experiment <- rbind(pretreat,treatment_range,posttreat)
experiment$treatpermil[is.na(experiment$treatpermil)]<-0
experiment$controlpermil[is.na(experiment$controlpermil)]<-0
experiment

ts.plot(experiment$treatpermil, col = "red")
ts.plot(experiment$controlpermil, col = "blue")

# Look at Raw Frequency Counts

treatment_ts <- ts(experiment$frequency.x, frequency = 12, start = c(2008,9))
control_ts <- ts(experiment$frequency.y, frequency = 12, start = c(2008,9))
ts.plot(control_ts, treatment_ts, main = "Breaches over time", xlim=c(2008,2011), gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Count per month")

# Look at Treatment and Control per Million

treatment_tsM <- ts(experiment$treatpermil, frequency = 12, start = c(2008,6))
mean(treatment_tsM)
sd(treatment_tsM)

control_tsM <- ts(experiment$controlpermil, frequency = 12, start = c(2008,6))
mean(control_tsM)
sd(control_tsM)

ts.plot(control_tsM, treatment_tsM, main = "Breaches over time", xlim=c(2008,2010), 
                    gpars = list(col = c("black", "red")), type = "b", xlab = "Years", ylab = "Count per month")


```




## Run Statistical Tests on Time Series for Stationarity
```{r}
# source of statistical tests http://r-statistics.co/Time-Series-Analysis-With-R.html

acfcontrol <- acf(control_ts) # autocorrelation (i.e. a Time Series with lags of itself)
acftreatment <- acf(treatment_ts)
# shows that the control time series is a "stationary time series"

png(here::here("Output","acfcontrol.png"))
plot(acfcontrol)

png(here::here("Output","acttreatmentNH.png"))
plot(acftreatment)

pacfcontrolMA <- pacf(control_ts)  # partial autocorrelation (i.e. correlation of the time series with a lag of itself, with the linear dependence of all the lags between them removed.)
pacftreatmentNH <- pacf(treatment_ts)  # partial autocorrelation (i.e. correlation of the time series with a lag of itself, with the linear dependence of all the lags between them removed.)

plot(pacfcontrolMA)

png(here::here("Output","pacftreatmentNH.png"))
plot(pacftreatmentNH)

ccfRes <- ccf(control_ts, treatment_ts, ylab = "cross-correlation")
ccfRes

# adf test is an Augmented Dickey-Fuller Test
adf.test(control_ts) # p-value < 0.05 indicates the TS is stationary
adf.test(treatment_ts)
kpss.test(control_ts) # Kwiatkowski-Phillips-Schmidt-Shin (KPSS) testz
kpss.test(treatment_ts)

# https://www.sas.com/content/dam/SAS/en_ca/User%20Group%20Presentations/Health-User-Groups/ITS_SAS.pdf

```


## ITS analyses use regression-based techniques
```{r}

quasiexp <- experiment[experiment$type != "test",]
quasiexp

# Added dummy variables for ITS
control <- as.data.frame(t(rbind(quasiexp$yearmonth,quasiexp$controlpermil)))
control$treat <- as.vector(rep(0,nrow(control)))                  # Create example vector
time <- 1:nrow(control)
control$time <- as.vector(time)
control$z <- c(rep(0,6),1:(nrow(control)-6))

treatment <- as.data.frame(t(rbind(quasiexp$yearmonth,quasiexp$treatpermil)))
treatment$treat <- as.vector(rep(1,nrow(control)))                  # Create example vector
time <- 1:nrow(control)
treatment$time <- as.vector(time)
treatment$z <- c(rep(0,6),1:(nrow(control)-6))
treatment

AppendITS <- rbind(treatment,control)
names(AppendITS) <- c("yearmonth","incident_permil","treat","time","z")
AppendITS$incident_permil <- as.numeric(as.character(AppendITS$incident_permil))
AppendITS$time <- as.numeric(as.character(AppendITS$time))
AppendITS$z <- as.numeric(as.character(AppendITS$z))
AppendITS

factor_cols <- c("treat","time","z")

sapply(AppendITS, class)

regTest <- lm(incident_permil ~ time + treat + z, AppendITS) 
summary(regTest)

regTest2 <- lm(incident_permil ~ time + treat + time*treat + z + z*time + z*treat + z*treat*time, AppendITS) 
summary(regTest2)

AppendITS

# View(AppendITS)

```