---
title: "Massachusetts_Case"
author: "Karl Grindal"
date: "6/8/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Case 1: Massachusetts Data Security Law 
```{r, Setup Chunk}
library(plyr)
library(here)
library(tidyr)
library(dplyr)
library(lubridate)
library(tseries)
library(TTR)
library(magrittr)

AllStateClean <- read.table(here::here("Data","Other_data","AllStateClean.txt"),sep=";")

AllStateClean <- AllStateClean %>% 
  mutate(MA_affected_residents = gsub("\\,.*","",MA_affected_residents)) %>% # this selects only the first number instance before a comma
  mutate(NC_affected_residents = gsub("\\,.*","",NC_affected_residents)) %>% # this selects only the first number instance before a comma
  mutate(Massachusetts = replace(Massachusetts, !is.na(Massachusetts), 1)) %>% 
  mutate(New_Hampshire = replace(New_Hampshire, !is.na(New_Hampshire), 1)) %>% 
  mutate(North_Carolina = replace(North_Carolina, !is.na(North_Carolina), 1)) %>% 
  replace_na(list(Massachusetts=0,New_Hampshire=0,North_Carolina=0)) 

```

# Identifying treatment and control options
```{r}
# Case 1: Experiment 1
massachusetts <- dplyr::filter(AllStateClean,Massachusetts==1)
massachusetts$reported_date <- substr(massachusetts$reported_date, start=1, stop=10)

new_hampshire <- dplyr::filter(AllStateClean,New_Hampshire==1)
new_hampshire$reported_date <- substr(new_hampshire$reported_date, start=1, stop=10)

# Case 1: Experiment 2
massachusetts$MA_affected_residents <- as.numeric(massachusetts$MA_affected_residents)
massachusetts1000 <- subset(massachusetts, massachusetts$MA_affected_residents > 1000)

n_carolina <- dplyr::filter(AllStateClean,North_Carolina==1)
n_carolina$NC_affected_residents <- as.numeric(n_carolina$NC_affected_residents)
n_carolina1000 <- subset(n_carolina, n_carolina$NC_affected_residents > 1000)

```

# Specifying Treatment and Control Variables
```{r}
Treat_Name <- 'Massachusetts'
Ctrl_Name <- 'North Carolina'

treatment <- massachusetts1000 # This Must Be Filled in to Work Properly!
control <- n_carolina1000  # This Must Be Filled in to Work Properly!
```

# Specifying Time Variables
```{r}
first_breach_date <- as.Date("2006-06-01")

pop_start <- c(2000,4)
# data_start <- c(2006,6) # New Hampshire
data_start	<- c(2005,12)	# North Carolina
data_range <- c(2006,2021)
exp_start <- c(2008,4)
exp_range <- c(2008,2011)
first_pop_date <- as.Date("2000-04-01")
last_pop_date <- as.Date("2020-12-01")
no_months_out <- 174

months_prior <- 5 # months before treatment start
months_after <- 5 # months after treatment start

# Legislation H.B. 4144 signed into law August 3, 2007
# Legislation H.B. 4144 becomes effective on October 31, 2007
# OCABR finalized the regulation on September 22, 2008

treatment_start <- as.Date("09/22/2008", "%m/%d/%Y") # Legislation H.B. 4144 becomes effective
treatment_end <- as.Date("03/01/2010", "%m/%d/%Y") # post 6 months after enforcement
```

# Produce Tables on Overlapping Co-occurance of Breach Incidents
```{r}
AllStateClean <- AllStateClean %>% 
    mutate(reported_date = substr(reported_date, start=1, stop=10)) %>%
    filter(reported_date > "2008-03-22" & reported_date < "2010-09-01") %>%
    mutate(MA_affected_residents = as.numeric(MA_affected_residents)) %>%
    mutate(NC_affected_residents = as.numeric(NC_affected_residents)) %>%
    mutate(MA_BigBreach = MA_affected_residents > 1000) %>%
    mutate(NC_BigBreach = NC_affected_residents > 1000)

AllStateClean

table(AllStateClean$Massachusetts) # 0 = 313, 1 = 858
table(AllStateClean$New_Hampshire) # 0 = 949, 1 = 222
table(AllStateClean$North_Carolina) # 0 = 819, 1 = 352

table(AllStateClean$Massachusetts,AllStateClean$New_Hampshire) # 0 = 205 + 108 ; 1 = 744 + 114
table(AllStateClean$Massachusetts,AllStateClean$North_Carolina) # 0 = 99 + 214 ; 1 = 720 + 138

# Need to resolve this R issue to get to a fixed solution
table(AllStateClean$MA_BigBreach,AllStateClean$NC_BigBreach)

```

# Create Population Time Series for Matching with Incident Frequncy
```{r}
# Population
pop <- read.csv(here::here("Data","Other_data","populations.csv")) # starts at 2000.04.01
datforpop <- data.frame(seq(first_pop_date, last_pop_date, by="1 month"))

datforpop <- datforpop %>%
  set_colnames(c("yearmonth")) %>%
  mutate(yearmonth=format(as.Date(yearmonth), "%Y/%m"))

pop <- pop %>%
  t() %>%
  as.data.frame() %>%
  set_colnames(pop[,1])

pop <- pop[-1,]
pop <- cbind(datforpop,pop)
rownames(pop) <- seq(1:nrow(pop))

freq_dat <- data.frame(seq(first_breach_date, last_pop_date, by="1 month"))
freq_dat <- freq_dat %>%
  set_colnames(c("yearmonth")) %>%
  mutate(yearmonth=format(as.Date(yearmonth), "%Y/%m"))

```


# Experiment 1: Create control and treatment populations with Total Incidents
```{r}

treatment_freq <- treatment %>% 
  dplyr::mutate(date_formatted=format(as.Date(reported_date), "%Y/%m")) %>% 
  dplyr::group_by(date_formatted)  %>% 
  dplyr::summarise(frequency = n()) %>% 
  set_colnames(c("yearmonth","frequency")) %>% 
  replace_na(list(frequency=0))

control_freq <- control %>% 
  dplyr::mutate(date_formatted=format(as.Date(reported_date), "%Y/%m")) %>% 
  dplyr::group_by(date_formatted)  %>% 
  dplyr::summarise(frequency = n()) %>% 
  set_colnames(c("yearmonth","frequency")) %>% 
  replace_na(list(frequency=0))

treatment_freq <- merge(treatment_freq,freq_dat, by="yearmonth", all.y=TRUE)
control_freq <- merge(control_freq,freq_dat, by="yearmonth", all.y=TRUE)
treatment_freq$frequency[is.na(treatment_freq$frequency)]<-0
control_freq$frequency[is.na(control_freq$frequency)]<-0

treatment_ts <- ts(treatment_freq$frequency, frequency = 12, start = data_start)
control_ts <- ts(control_freq$frequency, frequency = 12, start = data_start)

plot.ts(treatment_ts, main = "Breaches over time", xlim=data_range, xlab = "Years", ylab = "Count per month")
plot.ts(control_ts, main = "Breaches over time", xlim=data_range, xlab = "Years", ylab = "Count per month")

ts.plot(control_ts, treatment_ts, main = "Breaches over time", xlim=data_range, gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Count per month")

summary(treatment_ts)
summary(control_ts)

```


## Create charts with breaches per million residents
```{r}
# Merge Treatment and Control Together
comb_ts <- merge(treatment_freq, control_freq, by="yearmonth", all=TRUE)

# Merge Combined Treatment and Control Together with Population Statistics
comb_ts <- merge(comb_ts, pop, by='yearmonth', all.x = TRUE)
comb_ts[2:ncol(comb_ts)] <- sapply(comb_ts[2:ncol(comb_ts)],as.numeric)

comb_ts <- comb_ts %>% 
  replace_na(list(frequency.x=0,frequency.y=0)) %>% 
  dplyr::mutate(treatpermil = frequency.x/(Massachusetts/1000000)) %>% 
  dplyr::mutate(controlpermil = frequency.y/(`North Carolina`/1000000)) %>% 
  mutate(yearmonth2 = gsub("/","",comb_ts$yearmonth)) %>% 
  dplyr::filter(yearmonth2 > 200605) %>% 
  select(-("yearmonth2"))

treatment_tsM <- ts(comb_ts$treatpermil, frequency = 12, start = data_start)
control_tsM <- ts(comb_ts$controlpermil, frequency = 12, start = data_start)

plot.ts(treatment_tsM, main = "Breaches per Million over time", xlim=data_range, xlab = "Years", ylab = "Breaches per Million Resident During Month")
plot.ts(control_tsM, main = "Breaches per Million over time", xlim=data_range, xlab = "Years", ylab = "Breaches per Million Resident During Month")

ts.plot(control_tsM, treatment_tsM, main = "Breaches over time", xlim=data_range, gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Breaches per Million Resident During Month")
```

## Decompose the data to find seasonal patterns
```{r}
controltimeseriescomponents <- decompose(control_tsM)
plot(controltimeseriescomponents)
controltimeseriesseasonallyadjusted <- control_ts - controltimeseriescomponents$seasonal
plot(controltimeseriesseasonallyadjusted)

treatmenttimeseriescomponents <- decompose(treatment_tsM)
plot(treatmenttimeseriescomponents)
treatmenttimeseriesseasonallyadjusted <- treatment_ts - treatmenttimeseriescomponents$seasonal
plot(treatmenttimeseriesseasonallyadjusted)

ts.plot(controltimeseriesseasonallyadjusted, treatmenttimeseriesseasonallyadjusted, main = "Breaches over time", xlim=exp_range, gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Breaches per Million Resident During Month")

```

## Identifying and subsetting relevant dates
```{r}

treatment_start<- format(as.Date(as.character(treatment_start), origin = "1970-01-01"), "%Y/%m")
treatment_end<- format(as.Date(as.character(treatment_end), origin = "1970-01-01"), "%Y/%m")

pretreat <- comb_ts[(which(comb_ts$yearmonth==treatment_start)-months_prior):which(comb_ts$yearmonth==treatment_start),]
pretreat$type <- "pretest"

posttreat <- comb_ts[which(comb_ts$yearmonth==treatment_end):(which(comb_ts$yearmonth==treatment_end)+months_after),]
posttreat$type <- "posttest"

mean(posttreat$treatpermil) - mean(pretreat$treatpermil)
mean(posttreat$controlpermil) - mean(pretreat$controlpermil)

treatment_range <- comb_ts[(which(comb_ts$yearmonth==treatment_start)+1):(which(comb_ts$yearmonth==treatment_end)-1),]
treatment_range$type <- "test"

experiment <- rbind(pretreat,treatment_range,posttreat)
experiment$treatpermil[is.na(experiment$treatpermil)]<-0
experiment$controlpermil[is.na(experiment$controlpermil)]<-0

ts.plot(experiment$treatpermil, col = "red")
ts.plot(experiment$controlpermil, col = "blue")

# Look at Raw Frequency Counts

treatment_ts <- ts(experiment$frequency.x, frequency = 12, start = exp_start)
control_ts <- ts(experiment$frequency.y, frequency = 12, start = exp_start)
ts.plot(control_ts, treatment_ts, main = "Breaches over time", xlim=exp_range, gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Count per month")

# Look at Treatment and Control per Million

treatment_tsM <- ts(experiment$treatpermil, frequency = 12, start = exp_start)
mean(treatment_tsM)
sd(treatment_tsM)

control_tsM <- ts(experiment$controlpermil, frequency = 12, start = exp_start)
mean(control_tsM)
sd(control_tsM)

ts.plot(control_tsM, treatment_tsM, main = "Breaches over time", xlim=exp_range, 
                    gpars = list(col = c("black", "red")), type = "b", xlab = "Years", ylab = "Count per month")


```




## Run Statistical Tests on Time Series for Stationarity
```{r}
# source of statistical tests http://r-statistics.co/Time-Series-Analysis-With-R.html

acfcontrol <- acf(control_ts) # autocorrelation (i.e. a Time Series with lags of itself)
acftreatment <- acf(treatment_ts)
# shows that the control time series is a "stationary time series"

png(here::here("Output","acfcontrol.png"))
plot(acfcontrol)

png(here::here("Output","acftreatmentNH.png"))
plot(acftreatment)

pacfcontrolNH <- pacf(control_ts)  # partial autocorrelation (i.e. correlation of the time series with a lag of itself, with the linear dependence of all the lags between them removed.)
pacftreatmentMA <- pacf(treatment_ts)  # partial autocorrelation (i.e. correlation of the time series with a lag of itself, with the linear dependence of all the lags between them removed.)

png(here::here("Output","pacfcontrolNH.png"))
plot(pacfcontrolNH)

png(here::here("Output","pacftreatmentMA.png"))
plot(pacftreatmentMA)

ccfRes <- ccf(control_ts, treatment_ts, ylab = "cross-correlation")
ccfRes

# adf test is an Augmented Dickey-Fuller Test
adf.test(control_ts) # p-value < 0.05 indicates the TS is stationary
adf.test(treatment_ts)
kpss.test(control_ts) # Kwiatkowski-Phillips-Schmidt-Shin (KPSS) testz
kpss.test(treatment_ts)

# https://www.sas.com/content/dam/SAS/en_ca/User%20Group%20Presentations/Health-User-Groups/ITS_SAS.pdf

```


## ITS analyses use regression-based techniques
```{r}

quasiexp <- experiment[experiment$type != "test",]

# Added dummy variables for ITS
control <- as.data.frame(t(rbind(quasiexp$yearmonth,quasiexp$controlpermil)))
control$treat <- as.vector(rep(0,nrow(control)))                  # Create example vector
time <- 1:nrow(control)
control$time <- as.vector(time)
control$z <- c(rep(0,6),1:(nrow(control)-6))

treatment <- as.data.frame(t(rbind(quasiexp$yearmonth,quasiexp$treatpermil)))
treatment$treat <- as.vector(rep(1,nrow(control)))                  # Create example vector
time <- 1:nrow(control)
treatment$time <- as.vector(time)
treatment$z <- c(rep(0,6),1:(nrow(control)-6))
treatment

AppendITS <- rbind(treatment,control)
names(AppendITS) <- c("yearmonth","incident_permil","treat","time","z")
AppendITS$incident_permil <- as.numeric(as.character(AppendITS$incident_permil))
AppendITS$time <- as.numeric(as.character(AppendITS$time))
AppendITS$z <- as.numeric(as.character(AppendITS$z))
AppendITS

factor_cols <- c("treat","time","z")

sapply(AppendITS, class)

regTest <- lm(incident_permil ~ time + treat + z, AppendITS) 
summary(regTest)

regTest2 <- lm(incident_permil ~ time * treat * z, AppendITS) 
summary(regTest2)

AppendITS

# View(AppendITS)

```
