---
title: "Massachusetts_Case"
author: "Karl Grindal"
date: "6/8/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Case 1: Massachusetts Data Security Law 
```{r, Setup Chunk}
library(plyr)
library(here)
library(tidyr)
library(dplyr)
library(lubridate)
library(tseries)
library(TTR)

AllStateClean <- read.table(here::here("Data","Other_data","AllStateClean.txt"),sep=";")

AllStateClean$MA_affected_residents <- gsub("\\,.*","",AllStateClean$MA_affected_residents) # this selects only the first number instance before a comma
AllStateClean$NC_affected_residents <- gsub("\\,.*","",AllStateClean$NC_affected_residents) # this selects only the first number before a comma

AllStateClean$Massachusetts[!is.na(AllStateClean$Massachusetts)]<-1
AllStateClean$New_Hampshire[!is.na(AllStateClean$New_Hampshire)]<-1
AllStateClean$North_Carolina[!is.na(AllStateClean$North_Carolina)]<-1
AllStateClean$Massachusetts[is.na(AllStateClean$Massachusetts)]<-0
AllStateClean$New_Hampshire[is.na(AllStateClean$New_Hampshire)]<-0
AllStateClean$North_Carolina[is.na(AllStateClean$North_Carolina)]<-0

```

# Produce Tables on Overlapping Co-occurance of Breach Incidents
```{r}

CoveredDays <- AllStateClean

CoveredDays$reported_date <- substr(CoveredDays$reported_date, start=1, stop=10)

CoveredDays <- subset(CoveredDays, reported_date > as.Date("2008-03-22") )
CoveredDays <- subset(CoveredDays, reported_date < as.Date("2010-09-01") )

table(CoveredDays$Massachusetts) # 0 = 313, 1 = 858
table(CoveredDays$New_Hampshire) # 0 = 949, 1 = 222
table(CoveredDays$North_Carolina) # 0 = 819, 1 = 352

table(CoveredDays$Massachusetts,CoveredDays$New_Hampshire) # 0 = 205 + 108 ; 1 = 744 + 114
table(CoveredDays$Massachusetts,CoveredDays$North_Carolina) # 0 = 99 + 214 ; 1 = 720 + 138

CoveredDays$MA_affected_residents <- as.numeric(CoveredDays$MA_affected_residents)
CoveredDays$NC_affected_residents <- as.numeric(CoveredDays$NC_affected_residents)

CoveredDays$MA_BigBreach <- CoveredDays$MA_affected_residents > 1000
CoveredDays$NC_BigBreech <- CoveredDays$NC_affected_residents > 1000

# Need to resolve this R issue to get to a fixed solution

# table(CoveredDays$MA_BigBreach,CoveredDays$NC_BigBreach)

```

# Create Population Time Series for Matching with Incident Frequncy
```{r}
# Creating blank frequency starting with earliest date

dat2 <- data.frame(seq(as.Date("2006-06-01"), by="1 month", length.out=174)) # treatment date
names(dat2) <- "yearmonth"
dat2 <- format(dat2,  "%Y/%m")
dat2

# Population
pop <- read.csv(here::here("Data","Other_data","populations.csv")) # starts at 2000.04.01
pop <- pop[c(22,30,34),] # Massachusetts row 22, New Hampshire row 30, North Carolina row 34 # removes

datforpop <- data.frame(seq(as.Date("2000-04-01"), by="1 month", length.out=(length(pop)-1))) # treatment date
names(datforpop) <- "yearmonth"
datforpop <- format(datforpop,  "%Y/%m")
datforpop <- rbind("yearmonth",datforpop)
row.names(datforpop) <- 1:nrow(datforpop)

pop <- cbind(datforpop,t(pop))
colnames(pop) <- pop[1,]
pop <- pop[-1,]
rownames(pop) <- seq(1:nrow(pop))
pop <- as.data.frame(pop)
pop
```

# Identifying treatment and control options
```{r}
# Case 1: Experiment 1
massachusetts <- dplyr::filter(AllStateClean,Massachusetts==1)
massachusetts$reported_date <- substr(massachusetts$reported_date, start=1, stop=10)

new_hampshire <- dplyr::filter(AllStateClean,New_Hampshire==1)
new_hampshire$reported_date <- substr(new_hampshire$reported_date, start=1, stop=10)

# Case 1: Experiment 2
massachusetts <- dplyr::filter(AllStateClean,Massachusetts==1)
massachusetts$MA_affected_residents <- as.numeric(massachusetts$MA_affected_residents)
massachusetts1000 <- subset(massachusetts, massachusetts$MA_affected_residents > 1000)

n_carolina <- dplyr::filter(AllStateClean,North_Carolina==1)
n_carolina$NC_affected_residents <- as.numeric(n_carolina$NC_affected_residents)
n_carolina1000 <- subset(n_carolina, n_carolina$NC_affected_residents > 1000)

```

# Experiment 1: Create control and treatment populations (Massachusetts v New Hampshire) with Total Incidents
```{r}
treatment <- massachusetts1000 # This Must Be Filled in to Work Properly!
control <- n_carolina1000  # This Must Be Filled in to Work Properly!

# Format treatment dates into months
treatment$date_formatted <- format(as.Date(treatment$reported_date, "%Y-%m-%d"), "%Y/%m") # Alternative is "%m/%d/%Y"
treatment_freq <- treatment %>% 
  dplyr::group_by(treatment$date_formatted) %>% 
  dplyr::summarise(frequency = n(),)

names(treatment_freq)[1] <- "yearmonth"
treatment_freq$frequency[is.na(treatment_freq$frequency)]<-0

# Format control dates into months
control$date_formatted <- format(as.Date(control$reported_date, "%Y-%m-%d"), "%Y/%m") # Alternative is "%m/%d/%Y"
control_freq <- control %>% 
  dplyr::group_by(control$date_formatted) %>% 
  dplyr::summarise(frequency = n())

names(control_freq)[1] <- "yearmonth"
control_freq$frequency[is.na(control_freq$frequency)]<-0

treatment_ts <- ts(treatment_freq$frequency, frequency = 12, start = c(2006,6))
control_ts <- ts(control_freq$frequency, frequency = 12, start = c(2006,6))

plot.ts(treatment_ts, main = "Breaches over time", xlim=c(2006,2020), xlab = "Years", ylab = "Count per month")
plot.ts(control_ts, main = "Breaches over time", xlim=c(2006,2020), xlab = "Years", ylab = "Count per month")

ts.plot(control_ts, treatment_ts, main = "Breaches over time", xlim=c(2006,2020), gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Count per month")

summary(treatment_ts)
summary(control_ts)

```


## Create charts with breaches per million residents
```{r}
# Merge Treatment and Control Together
comb_ts <- merge(treatment_freq, control_freq, by="yearmonth", all=TRUE)

# Merge Combined Treatment and Control Together with Population Statistics
comb_ts <- merge(comb_ts, pop, by='yearmonth', all.x = TRUE)
comb_ts$frequency.x[is.na(comb_ts$frequency.x)]<-0
comb_ts$frequency.y[is.na(comb_ts$frequency.y)]<-0

comb_ts$Massachusetts <- as.numeric(as.character(comb_ts$Massachusetts))
comb_ts$treatpermil <- comb_ts$frequency.x/(comb_ts$Massachusetts/1000000)
class(comb_ts$frequency.y)<-"numeric"
comb_ts$`New Hampshire` <- as.numeric(as.character(comb_ts$`New Hampshire`))
comb_ts$controlpermil <- comb_ts$frequency.y/(comb_ts$`New Hampshire`/1000000)

treatment_tsM <- ts(comb_ts$treatpermil, frequency = 12, start = c(2006,6))
control_tsM <- ts(comb_ts$controlpermil, frequency = 12, start = c(2006,6))

plot.ts(treatment_tsM, main = "Breaches per Million over time", xlim=c(2006,2020), xlab = "Years", ylab = "Breaches per Million Resident During Month")
plot.ts(control_tsM, main = "Breaches per Million over time", xlim=c(2006,2020), xlab = "Years", ylab = "Breaches per Million Resident During Month")

ts.plot(control_tsM, treatment_tsM, main = "Breaches over time", xlim=c(2006,2020), gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Breaches per Million Resident During Month")
```

## Identifying and subsetting relevant dates
```{r}
# Legislation H.B. 4144 signed into law August 3, 2007
# Legislation H.B. 4144 becomes effective on October 31, 2007
# OCABR finalized the regulation on September 22, 2008

treatment_start <- as.Date("09/22/2008", "%m/%d/%Y") # Legislation H.B. 4144 becomes effective
treatment_start<- format(as.Date(as.character(treatment_start), origin = "1970-01-01"), "%Y/%m")

treatment_end <- as.Date("03/01/2010", "%m/%d/%Y")+180 # post 6 months after enforcement
treatment_end<- format(as.Date(as.character(treatment_end), origin = "1970-01-01"), "%Y/%m")

pretreat <- comb_ts[(which(comb_ts$yearmonth==treatment_start)-5):which(comb_ts$yearmonth==treatment_start),]
pretreat$type <- "pretest"

posttreat <- comb_ts[which(comb_ts$yearmonth==treatment_end):(which(comb_ts$yearmonth==treatment_end)+5),]
posttreat$type <- "posttest"

mean(posttreat$treatpermil) - mean(pretreat$treatpermil)
mean(posttreat$controlpermil) - mean(pretreat$controlpermil)

treatment_range <- comb_ts[(which(comb_ts$yearmonth==treatment_start)+1):(which(comb_ts$yearmonth==treatment_end)-1),]
treatment_range$type <- "test"

experiment <- rbind(pretreat,treatment_range,posttreat)
experiment$treatpermil[is.na(experiment$treatpermil)]<-0
experiment$controlpermil[is.na(experiment$controlpermil)]<-0
experiment

ts.plot(experiment$treatpermil, col = "red")
ts.plot(experiment$controlpermil, col = "blue")

# Look at Raw Frequency Counts

treatment_ts <- ts(experiment$frequency.x, frequency = 12, start = c(2008,4))
control_ts <- ts(experiment$frequency.y, frequency = 12, start = c(2008,4))
ts.plot(control_ts, treatment_ts, main = "Breaches over time", xlim=c(2008,2011), gpars = list(col = c("black", "red")), xlab = "Years", ylab = "Count per month")

# Look at Treatment and Control per Million

treatment_tsM <- ts(experiment$treatpermil, frequency = 12, start = c(2008,4))
mean(treatment_tsM)
sd(treatment_tsM)

control_tsM <- ts(experiment$controlpermil, frequency = 12, start = c(2008,4))
mean(control_tsM)
sd(control_tsM)

ts.plot(control_tsM, treatment_tsM, main = "Breaches over time", xlim=c(2008,2011), 
                    gpars = list(col = c("black", "red")), type = "b", xlab = "Years", ylab = "Count per month")


```

## Decompose the Massachusetts Data to Find Seasonal Patterns
```{r}

controltimeseriescomponents <- decompose(control_ts)
plot(controltimeseriescomponents)
controltimeseriesseasonallyadjusted <- control_ts - controltimeseriescomponents$seasonal
plot(controltimeseriesseasonallyadjusted)

```


## Run Statistical Tests on Time Series for Stationarity
```{r}
# source of statistical tests http://r-statistics.co/Time-Series-Analysis-With-R.html

acfcontrolMA <- acf(control_ts) # autocorrelation (i.e. a Time Series with lags of itself)
acttreatmentNH <- acf(treatment_ts)
# shows that the control time series is a "stationary time series"

png(here::here("Output","acfcontrolMA.png"))
plot(acfcontrolMA)

png(here::here("Output","acttreatmentNH.png"))
plot(acttreatmentNH)

pacfcontrolMA <- pacf(control_ts)  # partial autocorrelation (i.e. correlation of the time series with a lag of itself, with the linear dependence of all the lags between them removed.)
pacftreatmentNH <- pacf(treatment_ts)  # partial autocorrelation (i.e. correlation of the time series with a lag of itself, with the linear dependence of all the lags between them removed.)

png(here::here("Output","pacfcontrolMA.png"))
plot(pacfcontrolMA)

png(here::here("Output","pacftreatmentNH.png"))
plot(pacftreatmentNH)

ccfRes <- ccf(control_ts, treatment_ts, ylab = "cross-correlation")
ccfRes

# adf test is an Augmented Dickey-Fuller Test
adf.test(control_ts) # p-value < 0.05 indicates the TS is stationary
adf.test(treatment_ts)
kpss.test(control_ts) # Kwiatkowski-Phillips-Schmidt-Shin (KPSS) testz
kpss.test(treatment_ts)

# https://www.sas.com/content/dam/SAS/en_ca/User%20Group%20Presentations/Health-User-Groups/ITS_SAS.pdf

```


## ITS analyses use regression-based techniques
```{r}

quasiexp <- experiment[experiment$type != "test",]
quasiexp

# Added dummy variables for ITS
control <- as.data.frame(t(rbind(quasiexp$yearmonth,quasiexp$controlpermil)))
control$treat <- as.vector(rep(0,nrow(control)))                  # Create example vector
time <- 1:nrow(control)
control$time <- as.vector(time)
control$z <- c(rep(0,6),1:(nrow(control)-6))

treatment <- as.data.frame(t(rbind(quasiexp$yearmonth,quasiexp$treatpermil)))
treatment$treat <- as.vector(rep(1,nrow(control)))                  # Create example vector
time <- 1:nrow(control)
treatment$time <- as.vector(time)
treatment$z <- c(rep(0,6),1:(nrow(control)-6))
treatment

AppendITS <- rbind(treatment,control)
names(AppendITS) <- c("yearmonth","incident_permil","treat","time","z")
AppendITS$incident_permil <- as.numeric(as.character(AppendITS$incident_permil))
AppendITS$time <- as.numeric(as.character(AppendITS$time))
AppendITS$z <- as.numeric(as.character(AppendITS$z))
AppendITS

factor_cols <- c("treat","time","z")

sapply(AppendITS, class)

regTest <- lm(incident_permil ~ time + treat + z, AppendITS) 
summary(regTest)

regTest2 <- lm(incident_permil ~ time + treat + time*treat + z + z*time + z*treat + z*treat*time, AppendITS) 
summary(regTest2)

# View(AppendITS)

```
